Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-24 paper.qmd
Updated Figure 9 (c)

## To initialize benchmark values to remove low density hexagons
benchmark_rm_hex_vec <- append(seq(0, 0.99, by=0.1), c(benchmark1, 0.99))

error_rm_scurve <- data.frame(matrix(nrow = 0, ncol = 0))

@@

benchmark_label_df <- tibble(x = 0.2, y = 0.1, 
                          label = paste0("benchmark is ", benchmark1))

@@

mse_scurve_lwd <- ggplot(error_rm_scurve, 
                     aes(x = benchmark_rm_lwd, 
                         y = MSE)) + 
  geom_point(
    size = 1
    ) +
  geom_line(
    linewidth = 0.3
    ) + 
  # geom_vline(xintercept = benchmark1, linetype="solid", 
  #            color = "#bdbdbd", linewidth=1, alpha = 0.5) +
  # scale_x_continuous("standardized bin count", 
  #        transform = "reverse") + 
  scale_x_continuous(breaks = unique(error_rm_scurve$benchmark_rm_lwd)) +
  xlab("threshold to remove low-density hexagons") +
  ylab("MSE") +
  ggtitle("(c)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line())

# mse_scurve_lwd2 <- error_rm_scurve |>
#   filter(b_non_empty >= 5) |>
#   ggplot(aes(x = b_non_empty,y = MSE)) + 
#   geom_point(
#     size = 1
#     ) +
#   geom_line(
#     linewidth = 0.3
#     ) + 
#   # geom_vline(xintercept = benchmark1, linetype="solid", 
#   #            color = "#bdbdbd", linewidth=1, alpha = 0.5) +
#   scale_x_continuous("m (b1 = 14)") + 
#   ylab("MSE") +
#   ggtitle("(e)") +
#   theme(aspect.ratio = 0.75,
#         axis.text.x = element_text(size = 12),
#         axis.text.y = element_text(size = 12),
#         axis.title.x = element_text(size = 12),
#         axis.title.y = element_text(size = 12, 
#             angle=90),
#         plot.title = element_text(size = 12))
```

>>>>

## To initialize benchmark values to remove low density hexagons
benchmark_rm_hex_vec <- sort(append(seq(0, 0.6, by=0.15), benchmark1))

## option 1
error_rm_scurve <- data.frame(matrix(nrow = 0, ncol = 0))

@@

# benchmark_label_df <- tibble(x = 0.2, y = 0.1, 
#                           label = paste0("benchmark is ", benchmark1))

@@

### Option 2

scurve_model2 <- fit_highd_model(
  training_data = training_data_scurve,
  emb_df = umap_scurve_scaled,
  bin1 = 7,
  r2 = r2,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "x"
)

df_bin_centroids_scurve2 <- scurve_model2$df_bin_centroids
df_bin_scurve2 <- scurve_model2$df_bin

error_rm_scurve2 <- data.frame(matrix(nrow = 0, ncol = 0))

  df_bin_centroids_scurve_high_dens <- df_bin_centroids_scurve2 |>
    filter(std_counts > benchmark_rm_lwd)

  df_bin_scurve_high_dens <- df_bin_scurve2 |>
    filter(hb_id %in% df_bin_centroids_scurve_high_dens$hexID)

  ## Compute error
  error_df <- glance(
    df_bin_centroids = df_bin_centroids_scurve_high_dens,
    df_bin = df_bin_scurve_high_dens,
    training_data = training_data_scurve,
    newdata = NULL,
    type_NLDR = "UMAP",
    col_start = "x") |>
    mutate(benchmark_rm_lwd = round(benchmark_rm_lwd, 2),
           bin1 = 7,
           bin2 = 13,
           b = bin1 * bin2,
           b_non_empty = NROW(df_bin_centroids_scurve_high_dens),
           mean_counts = sum(df_bin_centroids_scurve_high_dens$bin_counts)/NROW(df_bin_centroids_scurve_high_dens))

  error_rm_scurve2 <- bind_rows(error_rm_scurve2, error_df)

### Option 3

scurve_model3 <- fit_highd_model(
  training_data = training_data_scurve,
  emb_df = umap_scurve_scaled,
  bin1 = 13,
  r2 = r2,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "x"
)

df_bin_centroids_scurve3 <- scurve_model3$df_bin_centroids
df_bin_scurve3 <- scurve_model3$df_bin

error_rm_scurve3 <- data.frame(matrix(nrow = 0, ncol = 0))

  df_bin_centroids_scurve_high_dens <- df_bin_centroids_scurve3 |>
    filter(std_counts > benchmark_rm_lwd)

  df_bin_scurve_high_dens <- df_bin_scurve3 |>
    filter(hb_id %in% df_bin_centroids_scurve_high_dens$hexID)

  ## Compute error
  error_df <- glance(
    df_bin_centroids = df_bin_centroids_scurve_high_dens,
    df_bin = df_bin_scurve_high_dens,
    training_data = training_data_scurve,
    newdata = NULL,
    type_NLDR = "UMAP",
    col_start = "x") |>
    mutate(benchmark_rm_lwd = round(benchmark_rm_lwd, 2),
           bin1 = 13,
           bin2 = 24,
           b = bin1 * bin2,
           b_non_empty = NROW(df_bin_centroids_scurve_high_dens),
           mean_counts = sum(df_bin_centroids_scurve_high_dens$bin_counts)/NROW(df_bin_centroids_scurve_high_dens))

  error_rm_scurve3 <- bind_rows(error_rm_scurve3, error_df)

error_rm_scurve <- bind_rows(
  error_rm_scurve,
  error_rm_scurve2,
  error_rm_scurve3
) |>
  mutate(bin1  = as.factor(bin1))

mse_scurve_lwd <- ggplot(error_rm_scurve, 
                     aes(x = benchmark_rm_lwd, 
                         y = MSE,
                         color = bin1)) + 
  geom_point(
    size = 1
    ) +
  geom_line(
    linewidth = 0.3
    ) + 
  # geom_vline(xintercept = benchmark1, linetype="solid", 
  #            color = "#bdbdbd", linewidth=1, alpha = 0.5) +
  # scale_x_continuous("standardized bin count", 
  #        transform = "reverse") + 
  scale_x_continuous(breaks = unique(error_rm_scurve$benchmark_rm_lwd)) +
  scale_color_manual(values=c("#08519c", "#006d2c", "#54278f")) +
  xlab("threshold to remove low-density hexagons") +
  ylab("MSE") +
  ggtitle("(c)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        legend.position = "none")
```
